\documentclass[a4paper,11pt]{scrartcl}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}


%% pacchetti 
%% ------------------------------------------------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{lmodern, fancyhdr, lastpage, xspace, marvosym }
\usepackage[margin=25mm,
			%includehead
			]{geometry}
\usepackage{setspace}
\usepackage{lscape}
\usepackage{clipboard} %for copying text in several parts of the document






\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{csquotes}
\usepackage[nohyperlinks,nolist]{acronym}

\usepackage{enumitem}

\usepackage{soul} % for smarter (word-wrapping) underlining.
\usepackage{tcolorbox}

\usepackage[textsize=9]{todonotes}
\newcommand{\todocd}[1]{{\textcolor{orange}{#1}}}
\newcommand{\todohost}[1]{{\textcolor{green}{#1}}}
\newcommand{\ignore}[1]{}

%tables & pictures
%------------------------------------------------------------------------------------
\usepackage{multirow,bigdelim, caption, longtable} % longtable for tables that can fit on multiple pages
\captionsetup[longtable]{position=below} 
\usepackage{booktabs}
\usepackage{graphicx}

%% bibliography
%%--------------------------------------------------------------------------------

% usata con \nobibliography
\usepackage[ibidem, titleformat=commasep, lookat, citefull=first]{jurabib}
\renewcommand{\cite}{\footcite} % citations in footnotes




%% math
%%------------------------------------------------------------------------------------
\usepackage{amsfonts, amsmath, amsthm, amssymb, mathrsfs, amscd}		
\usepackage{xfrac}

%% hyperref e pacchetti fastidiosi
%%--------------------------------------------------------------------------------
\usepackage{xurl}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}


%% ridefinizioni e layout
%% -------------------------------------------------------------------------------
% \usepackage[subtle, mathdisplays=tight, lists=tight, bibnotes=tight, indent=tight,  
%% sections=tight,
%%charwidths=tight
%  ]{savetrees}
%  

%\headheight=14pt

%% Explicitly set footnote font size to match call (i.e., 8pt).
%% Taken from http://tex.stackexchange.com/a/249422/84485
\makeatletter
\renewcommand\footnotesize{%
	\@setfontsize\footnotesize\@ixpt{8}%
	\abovedisplayskip 8\p@ \@plus2\p@ \@minus4\p@
	\abovedisplayshortskip \z@ \@plus\p@
	\belowdisplayshortskip 4\p@ \@plus2\p@ \@minus2\p@
	\def\@listi{\leftmargin\leftmargini
		\topsep 4\p@ \@plus2\p@ \@minus2\p@
		\parsep 2\p@ \@plus\p@ \@minus\p@
		\itemsep \parsep}%
	\belowdisplayskip \abovedisplayskip
}
\makeatother







%% To correctly align fancy headers.
%% Courtesy of: http://tex.stackexchange.com/a/88136/84485
\makeatletter
\newcommand{\resetHeadWidth}{\fancy@setoffs}
\makeatother


\pagestyle{fancy}
\fancyhead{}
\fancyhead[R]{\color{gray}{Page \thepage~of \pageref*{sec:endpage}}}
\fancyhead[C]{\color{gray}{Small groups activity, \today}}
\fancyhead[L]{\color{gray}{Clinical studies}}
\cfoot{}



%%definizioni per mate
%%------------------------------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]

\theoremstyle{remark}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{propy}[thm]{Property}
\newtheorem{step}[thm]{Step}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{exam}[thm]{Example}
\newtheorem{note}[thm]{Note}
\newtheorem{quest}[thm]{Question}



%%tikz stuff 
%%-----------------------------------------------------------------------------------
\usetikzlibrary{positioning,calc} 
\tikzset{shifted by/.style={to path={($(\tikztostart)!#1!90:(\tikztotarget)$)
 -- ($(\tikztotarget)!#1!-90:(\tikztostart)$)}}, 
 shifted by/.default=2pt,standard edge/.style={very thick,-latex}, 
 back and forth between/.style args={#1 and #2}{insert path={
  #1 edge[standard edge,-latex,shifted by] #2 #2 edge[standard edge,shifted by] #1}}} 
  
  
  
%% global settings
%%-----------------------------------------------------------------------------------


\title{Designing a clinical study for pancreatic cancer risk factors}
\date{}


%%-----------------------------------------------------------------------------------
%% begin document
%%-----------------------------------------------------------------------------------

\begin{document}
\begin{center}
\doublespacing
{\LARGE DESIGNING A CLINICAL STUDY FOR PANCREATIC CANCER RISK FACTORS}
\end{center}


%% In the following: instructions for the tutor. Note that
%% what is written in \Copy{scenario}{ - } is authomatically repeated
%% in the students' materials
%%-----------------------------------------------------------------------------------

\section{Introduction}
\label{sec:introduction}
\ignore{
The purpose of this activity is to give students the opportunity to think about the issues involved in designing case-control and cohort studies (future development: think about adding a cross-sectional study and a randomised control trial). The activity is designed to be used both for face-to-face teaching and for distant learners, with the due adaptations. A sheet will be handed out with the scenario and the scheme here included, and another, pre-filled with answers will be made available at the end of the activity. 

   
\section{Presentation}

\textit{\textbf{Don't hand out sheets yet. }}

\medskip

With the use of a presentation, recap case-control and cohort studies designs, what is a biobank. Then move to the questions here below. 


\section{Setting the scenario}

\textit{\textbf{Have a summary on the presentation and read out the following. This text will also be on the sheets that students will be handed later. }}
\medskip

}
Pancreatic cancer is the 10th most common cancer with 10,449 people diagnosed with pancreatic cancer in the UK in 2018. It has the lowest survival of all common cancers, with five-year survival less than 7\%. 

\medskip

Early diagnosis is crucial to improve survival outcomes, with one-year survival in those diagnosed at an early stage six times higher than one-year survival in those diagnosed at stage four.  
However, most people with pancreatic cancer are diagnosed at a late stage (UK statistics on \url{https://www.pancreaticcancer.org.uk/}). 
\medskip

Patients usually experience no specific symptoms until the cancer is in a late stage. Currently there is no screening programme for pancreatic cancer in the UK, nor are there data to show reduced mortality from pancreatic cancer from any screening test. The diagnosis is usually made by imaging, after late symptoms such as jaundice and weight loss have already developed. Some data provide preliminary support for use of imaging methods (endoscopic ultrasound and MRI) if patients screened are \textbf{at increased risk} of pancreatic cancer. Therefore an accurate, non-invasive way to identify people at sufficiently high risk could enable the development of pancreatic cancer screening. 

\medskip

Known risk factors for pancreatic cancers include age, sex, smoking, high body mass index (BMI),  diabetes, certain chemicals, ethnicity, and family history. There have also been investigations on the utility of a variety of \textbf{biomarkers} to identify those at sufficiently increased risk of pancreatic cancer to justify screening using imaging.

\medskip
Cohort studies have a long history, but a recent development has been the use of biobanks from large population cohorts (such as UK Biobank - \url{https://www.ukbiobank.ac.uk}), that record data on participants from traditional questionnaires, in addition to biological samples, for instance to help evaluate the association between genetic variation, environmental exposures for risk of disease. Disease-specific biobanks have also been developed. One such at Barts Cancer Institute is the Pancreatic Cancer Research Fund Tissue Bank (PCRF, for further details see \url{https://www.thepancreastissuebank.org}). Set up in 2016, it has collected:
\begin{itemize}
\item Blood samples from 2,200 consenting patients who underwent biopsies or surgery for pancreatic diseases, including pancreatic cancers (also urine, saliva and tissue samples);
\item Large numbers of healthy control blood (also urine and saliva).
\end{itemize}

\ignore{

\subsection{Open questions}
\textit{\textbf{Ask, and take notes of answers on the whiteboard:}}

\begin{quest} ``How would you design a study to investigate whether or not such a proposed biomarker (or panel of biomarkers) is an effective tool for identifying those at high risk of pancreatic cancer, in order to enable early detection of pancreatic cancer?'' 
\end{quest}


\begin{itshape}
Hopefully students will get to (after prompting if necessary):
\begin{itemize}
\item \textbf{Cross-sectional} study: test everyone that goes to a GP appointment (or better, target a high risk group somehow – eg. If registry people with strong family history for pancreatic cancer, or patients with diabetes mellitus) and record if they have a diagnosis of pancreatic cancers. Weak, no long term outcome; and pancreatic cancer is very rare so study would have to be enormous to find even a handful of cases.
\item Could do a \textbf{case-control} study to assess the performance of the biomarker in detecting pancreatic cancers across all stages. Any of the following might be discussed (don’t need to go through them all if not though) 
	\begin{itemize}
	\item Collect samples from patients with a new diagnosis of pancreatic cancer (blood and tissue) and controls without a diagnosis of cancer (blood). Think about who controls will be recruited and how.
	\item Use a biobank focused on pancreatic cancer for selecting participants with cancer, and controls without cancer (e.g. the Pancreatic Cancer Research Fund Tissue mentioned above \url{https://www.thepancreastissuebank.org}). Or use a biobank of general population (e.g. UK Biobank~\url{https://www.ukbiobank.ac.uk}) where samples taken at entry and stored at low temperature. Do a \textbf{nested case-control} study by taking cases diagnosed within a certain time of entry, and matched controls, and use the biobank samples to run the tests. Suggest that we are working under the assumption that technology works on the available material and that storage does not degrade samples. 
	\end{itemize}
\item Do a \textbf{cohort} study to validate the use of the test in detecting pancreatic cancers earlier. Remark that the number of participants would be huge, unless a high risk group may be identified (e.g. diabetes mellitus, strong family history, ...). Think about possible ethical issues.
\item Could do a \textbf{randomised controlled trial}, where we use the biomarker to take some sort of action (screening with imaging, clinical work-up), but also here one must be careful of ethical issues (e.g. participants whose tests detects potential signals of cancer in their blood, or seem at very high risk, will need to be told their group and be referred for further tests).
\end{itemize}
\end{itshape}

\begin{note}
If students identify cohort and case-control studies as \textbf{prospective} and \textbf{retrospective} studies, try and discourage these terms. Cohort and case-control are more accurate, especially as you can do a prospective case-control study (nested case-control – where at the study's end, the cases and controls are sampled from a cohort) and you can do a retrospective cohort study where you collect a cohort from historical data.
\end{note}

\textit{\textbf{Ask, and take notes of answers on the whiteboard:}}

\begin{quest} ``What summary measure of risk associated with the biomarker could you use for each study?''
\end{quest}

\begin{itshape}
Classical metrics:
\begin{itemize}
\item Cohort study: relative risk – for example risk of pancreatic cancer in people who were biomarker positive divided by risk  in people who were biomarker negative, if binary. If continuous: relative risk per unit of the biomarker, or per SD of the biomarker. Also (but not strictly relevant to the question here discussed): hazard ratio, and time-to-event (where the event can be the diagnosis).
\item Case-control study:  odds ratio (again could be binary or continuous per unit etc).
\end{itemize}

Metrics for diagnostic tests (not strictly relevant to the question asked, but good to discuss):
\begin{itemize}
\item Sensitivity: the proportion of test-positive subjects out of all disease-positive subjects. Independent of prevalence.
\item Specificity: proportion of test-negative subjects out of all disease-negative subjects. Independent of prevalence.
\item Positive Predictive Values (PPV): probability that following a positive test result, that individual will truly have that specific disease. Depends on prevalence. 
\item Negative Predictive Values (NPV): probability that following a negative test result, that individual will truly not have that specific disease. Depends on prevalence. 
\end{itemize}

\textbf{Summary:} either take people without disease but where some have and some don’t have the risk factor (in this case biomarker positive) and follow them up to assess risk in each group, hence RR; or take people with disease and without disease, and assess them for the presence of the risk factor (biomarker positive, hence odds ratio. Remark that it is equivalent to RR it the disease is rare.

\end{itshape}
}
\section{Group activity}
\ignore{
\subsection{Work in small groups (30 minutes)}

\textit{\textbf{Split the students into (muliples of) two groups (no more than 8 in each group) and tell them:}}

\medskip

}
Assume that Cancer Research UK is going to fund two different studies to investigate whether your research group’s biomarker panel can help stratify risk of pancreatic cancer, and enable early detection that might lead to improvements in public health outcomes. One will be a case-control study, that will report its results within 3 years and the second group they need to design a cohort study which will report its results within 13 years.

\medskip


\begin{description}
\item[Group A (and multiples)] needs to design a case-control study that reports results within 3 years;
\item[Group B] design a cohort study that reports results within 13 years.
\end{description}

\ignore{
\textit{\textbf{Hand out the student sheets which contain the questions to fill in to design the studies (the first group should answer the left hand column and the second group the right hand column).}}



\subsection{Group presentations (25 minutes)}


%%-----------------------------------------------------
%% Students materials 
%%-----------------------------------------------------

}
\begin{landscape}
\section{Annotated students' materials}
\subsection{Cohort and case control studies and possible metrics}

%changing font size for the students materials to be clearer
\begin{large}

\begin{center}
\LARGE{Cohort study}
\end{center}

\begin{center}
\begin{tikzpicture}[boot/.style={draw,thick,align=center, inner sep=2mm}]
\node[boot=20] (CH) {Identify the cohort};
\node[above right=0.5cm and 1.5cm of CH.east,boot=20] (EP) {Exposed people};
\node[below right=0.5cm and 1.5cm of CH.east,boot=20] (UP) {Unexposed people};
\node[right=1.5cm of EP,boot=20] (ID1) {Incidence of disease};
\node[right=1.5cm of UP, below=1cm of ID1, boot=20] (ID2) {Incidence of disease};
\node[right=1.5cm of ID1,boot=20] (RE) {Risk if exposed $= R_{ex}$};
\node[right=1.5cm of ID2, below=1cm of RE, boot=20] (RU) {Risk if unexposed $= R_{u}$};
\node[right=16cm of CH,boot=20] (STAT) {Relative Risk $= R_{ex} / R_{u}$ \\ Difference in risk $= R_{ex} - R_{u}$ \\ Hazard ratio \\ Time-to-event};

\node[below=2cm of CH,align=center] (SS) {Start of study};
\node[right=1cm of SS, below=1cm of UP,align=center] (ME) {Measure exposure and follow-up};
\node[right = 1cm of ME, align=center] (OC) {Outcome};
\node[right=18cm of SS,align=center] (STATD) {Statistics};

\draw (CH) edge[standard edge] (EP)
(CH) edge[standard edge] (UP)
(CH) edge[standard edge] (EP)
(EP) edge[standard edge] (ID1)
(UP) edge[standard edge] (ID2)
(ID1) edge[standard edge] (RE)
(ID2) edge[standard edge] (RU)
(RE) edge[standard edge] (STAT)
(RU) edge[standard edge] (STAT)

;
\end{tikzpicture}
\end{center}




\begin{center}
\LARGE{Case-control study}
\end{center}

\begin{center}
\begin{tikzpicture}[boot/.style={draw,thick,align=center, inner sep=2mm}]

\node[boot=20] (STAT) {Odds Ratio of disease \\  for exposed and not $ = O_{d^+}/ O_{d^-}$\\ Diagnostic test metrics };
\node[above right=0.5cm and 2cm of STAT.east,boot=20] (OD) {Odds of exposure if \\ disease $=O_{d^+}$};
\node[below right=0.5cm and 2cm of STAT.east,boot=20] (OND) {Odds of exposure if \\ no disease $=O_{d^-}$};
\node[right=2cm of OD,boot=20] (ME1) {Measure exposure};
\node[right=2cm of OND, boot=20] (ME2) {Measure exposure};
\node[right=2cm of ME1,boot=20, anchor=west] (ICa) {Identify cases with disease};
\node[right=2cm of ME2, boot=20, anchor=west] (ICo) {Identify controls without disease};



\node[below=1cm of ICo,align=center] (SS) {Start of study};
\node[left=3.2cm of SS, align=center] (IPE) {Investigate prior exposure};
\node[left=16.5cm of SS,align=center] (STATD) {Statistics};

\draw (ICa) edge[standard edge] (ME1)
(ICo) edge[standard edge] (ME2)
(ME1) edge[standard edge] (OD)
(ME2) edge[standard edge] (OND)
(OD) edge[standard edge] (STAT)
(OND) edge[standard edge] (STAT);
\end{tikzpicture}
\end{center}


\ignore{

\clearpage

\Paste{scenario}

\bigskip

\Paste{TheQuestion}

\bigskip

In answering the question you will need to think about the following. 
}

\clearpage

\begin{longtable}{p{0.45\linewidth} | p{0.45\linewidth}  }
%% Title row -------------------------------
\toprule
\textbf{A: Case-control study with results in 3 years} &\textbf{B: Cohort study with results in 13 years} \\ 
\toprule

%% First question -------------------------------
{\raggedright{{What are your entry criteria for \textbf{cases}?}}}
\vspace{2.5cm}
&
\multirow{3}{=}{
\raggedright{What characteristics of the cohort will make the follow up easier and enable you to obtain useful results? Think about who will you recruit - what type of people, for where and how, how many? }}
\vspace{2.5cm}
\\

\cmidrule{1-1} & \\
%% Second question -------------------------------
\raggedright{Most case-control studies match cases and controls. What would you match on and what are your entry criteria for \textbf{controls}?}
\vspace{2.5cm}
& \\
\cmidrule{1-1} & \\
%% Third question -------------------------------
\raggedright{Describe how you will recruit to the case control study (remember you can use a biobank)}
\vspace{2.5cm}
& \\
\pagebreak
\midrule
%% Forth question -------------------------------
\multicolumn{2}{c}{What is a simple binary risk factor or exposure?} \\ 
\vspace{2.5cm}
&
\vspace{2.5cm}
\\
\midrule
%% Question -------------------------------
\multicolumn{2}{c}{How else could you measure the risk factor (biomarker)?} \\ 
\multicolumn{2}{c} {
\vspace{2.5cm}}
 \\
\midrule
%% Question -------------------------------
\multicolumn{2}{c}{Briefly describe how you would collect information on the risk factor.} \\ 
\vspace{2.5cm}
 & 
\vspace{2.5cm}
\\
\midrule
%% Question -------------------------------
\multicolumn{2}{c}{Give the most accurate and useful simple binary measure of outcome for your study.} \\ 
\vspace{2.5cm}
&
\vspace{2.5cm}
 \\
%% Question -------------------------------
\pagebreak
\midrule
\multicolumn{2}{c}{Who will the actual results be representative of?} \\ 
\vspace{2.5cm}
& 
\vspace{2.5cm}
\\
\midrule
%% Question -------------------------------
\multicolumn{2}{c}{Who will these results be useful for?} \\ 
\multicolumn{2}{c}{
\vspace{2.5cm}}
 \\
\midrule
%% Question -------------------------------
\multicolumn{2}{c}{What other results could your get from your study group?} \\
\vspace{2.5cm}
&
\vspace{2.5cm}
\\
\midrule
%% Question----------------------------------
\multicolumn{2}{c}{Could the age of the participants affect the interpretation of your results?  If so how?} \\ 
\vspace{2.5cm}
&
\vspace{2.5cm}
\\
\pagebreak
\midrule
\multicolumn{2}{c}{What biases could occur from your study design?} \\ 
\vspace{2.5cm}
& 
\vspace{2.5cm}
\\
\midrule
%% Question ------------------------------------
\multicolumn{2}{p{0.9\linewidth}}
{\centering Give an example of a study of an exposure and a disease that would be unsuitable for your type of study but suitable for the other.\newline Explain the characteristics of the exposure or disease that make it unsuitable} \\ 
\vspace{2.5cm}
 & 
\vspace{2.5cm}
\\
\midrule
%% Question ---------------------------------
\multicolumn{2}{c}{Summarise the advantages in general of your type of study?} \\ 
\vspace{2.5cm}
&
\vspace{2.5cm}

 \\
\midrule
%% Question ------------------------------
\multicolumn{2}{c}{Summarise the disadvantages in general of your type of study?} \\ 
\vspace{2.5cm}
&
\vspace{2.5cm}
 \\
\bottomrule 
\caption{Fill in the side corresponding to your group.}
\label{T:twostudies}
\end{longtable}
 
Further reading:
\begin{itemize}
\item Early detection of pancreatic cancer: \url{https://pubmed.ncbi.nlm.nih.gov/32135127/}
\item Description of the UROPANC trial: \url{https://www.bartscancer.london/centre-for-experimental-cancer-medicine/clinical-trial-portfolio/uropanc/}
\item A combination of urinary biomarker panel and PancRISK score for earlier detection of pancreatic cancer: A case–control study: \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7758047/}
\item Germline BRCA2 K3326X and CHEK2 I157T mutations increase risk for sporadic pancreatic ductal adenocarcinoma: \url{https://onlinelibrary.wiley.com/doi/10.1002/ijc.32127}
\item Noninvasive Diagnosis of Pancreatic Cancer Through Detection of Volatile Organic Compounds in Urine: \url{https://pubmed.ncbi.nlm.nih.gov/29129714/}
\item Biomarkers in molecular medicine: cancer detection and diagnosis: \url{https://www.future-science.com/doi/10.2144/05384SU04}
\item Empirical Evidence of Design-Related Bias in Studies of Diagnostic Tests: \url{https://jamanetwork.com/journals/jama/fullarticle/191668}
\item Statistical Hypothesis Testing versus Machine Learning Binary Classification: Distinctions and Guidelines: \url{https://www.sciencedirect.com/science/article/pii/S2666389920301562}
\end{itemize}



\end{large}
\label{sec:endpage}
\end{landscape}
\end{document}
